**LCA(Latent Class Analysis) with variables age, admission type (elective or emergency) and morbiditiy composition of each patient**
```{r}
library(RPostgreSQL)
library(poLCA)
```
Establish connection to database in PostgreSQL
```{r}
#Establish connection to database in PostgreSQL
conn <- dbConnect(drv="PostgreSQL",
                  dbname="mimiciii",
                  host= "localhost",
                  user="postgres")
```
Query* to select necessary variables -

*Assuming create_elixhauser, create_patient_info,
*and create_elixhauser_final have been run in database already,
```{r}
# *Assuming create_elixhauser and #create_patient_info 
# *and create_elixhauser_final
# *have been run in database already
# Query* to select elixhauser_final -

query <- "SELECT age, admission_type_encoded, congestive_heart_failure, 
                     cardiac_arrhythmias, valvular_disease, 
                     pulmonary_circulation,
                     peripheral_vascular, hypertension,
                     paralysis,
                     other_neurological,
                     chronic_pulmonary,
                     diabetes_uncomplicated,
                     diabetes_complicated, hypothyroidism,
                     renal_failure,
                     liver_disease,
                     peptic_ulcer,
                     aids,
                     lymphoma,
                     solid_tumor,
                     rheumatoid_arthritis,
                     coagulopathy,
                     obesity,
                     weight_loss,
                     fluid_electrolyte,
                     blood_loss_anemia,
                     deficiency_anemias,
                     alcohol_abuse,
                     drug_abuse,
                     psychoses,
                     depression FROM ELIXHAUSER_FINAL;"
```
Store query results

```{r}
final_table <- dbGetQuery(conn, query)
```

Factor all variables for LCA (poLCA requires factored data)
```{r}
final_table$age <- as.factor(final_table$age)
final_table$admission_type_encoded <- as.factor(final_table$admission_type_encoded)
comorbidity_columns <- c(
  "congestive_heart_failure",
  "cardiac_arrhythmias",
  "valvular_disease",
  "pulmonary_circulation",
  "peripheral_vascular",
  "hypertension",
  "paralysis",
  "other_neurological",
  "chronic_pulmonary",
  "diabetes_uncomplicated",
  "diabetes_complicated",
  "hypothyroidism",
  "renal_failure",
  "liver_disease",
  "peptic_ulcer",
  "aids",
  "lymphoma",
  "solid_tumor",
  "rheumatoid_arthritis",
  "coagulopathy",
  "obesity",
  "weight_loss",
  "fluid_electrolyte",
  "blood_loss_anemia",
  "deficiency_anemias",
  "alcohol_abuse",
  "drug_abuse",
  "psychoses",
  "depression"
)

final_table$comorbidity_columns <- as.factor(final_table$comorbidity_columns)

final_table[comorbidity_columns] <- lapply(
                              final_table[comorbidity_columns], as.factor)
```

Define formula for running LCA with poLCA, cbind creates matrix 
with observed variables, and "~1" indicates 1 latent class being created
```{r}
lca_formula <- cbind(age, admission_type_encoded, congestive_heart_failure, 
                     cardiac_arrhythmias, valvular_disease, 
                     pulmonary_circulation,
                     peripheral_vascular, hypertension,
                     paralysis,
                     other_neurological,
                     chronic_pulmonary,
                     diabetes_uncomplicated,
                     diabetes_complicated, hypothyroidism,
                     renal_failure,
                     liver_disease,
                     peptic_ulcer,
                     aids,
                     lymphoma,
                     solid_tumor,
                     rheumatoid_arthritis,
                     coagulopathy,
                     obesity,
                     weight_loss,
                     fluid_electrolyte,
                     blood_loss_anemia,
                     deficiency_anemias,
                     alcohol_abuse,
                     drug_abuse,
                     psychoses,
                     depression) ~1

```

Run LCA model, we are using 5 Latent classes to model the data here
```{r}
lca_out <- poLCA(lca_formula, data=final_table, nclass=5, graph=TRUE)

dbDisconnect(conn)
```
